---
title: "The FEMSdevPkg"
output: 
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 2

vignette: >
  %\VignetteIndexEntry{FEMSdevPkg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: "Henriette Elise Breymann and others"
date: "`r Sys.Date()`"
---

\newpage


# Introduction

## The digital financial world

Today's financial system is digital and banks can rightly be considered as 
applied IT companys. Our money, our investments and debts are just zillions of 
numbers in
huge data bases. From this persective the banking business can be
considered of a transformation of these numbers according to appropriate rules.
And these transformations must be highly automated; otherwise banks could not function. 

Unfortunately, the automation does not work totally automated across the boards. 
Sure, the parts that are vital for banks to subsist do work automatically. 
This mainly concerns financial transactions.
But there are other parts that, even though equally important for controlling 
and steering the institutions, are
not as vital for the daily business and thus less integrated. 
Which generates
large costs for financial institutions and limits the impact of what could be 
achieved with optimal automation.

The technical reason is the fragmented IT landscape which itself is a result of 
the way this
landscape grew over the years, piece by piece. Each piece --- or subsystem ---
works well for itself but they are not optimally working together. The problem is
the interfaces and, closely related to this, the data standards. And these two very
technical aspects are closely related to the task that they should support, namely
banking. Which means that optimal interfaces and data standards that are up to the
task cannot be constructed without a thorough understanding of banking. An here is
the rub: IT and banking expertise are rarely found together in a single person.
Bankers care little about the underlying IT and IT specialists usually are not very 
literate in banking. Moreover, different lines of business large work independent
from one another. Which leads to silo-type structures: integration well advanced
within a silo but pur across silos. Which means that integration across silos
still requires a lot of manual work. As a consequence, regulatory reporting, which 
requires presicely this integration, is slow and costly. And the cost-income ratio
of banks is around 60\% or higher. 


To overcome these shortcoming, 
a special infrastructural layer has been developed. 
The layer we're refering to is the ACTUS 
standard for digital financial contract representation. ACTUS stands for 
Algorithmic Contract Type Unified Standard. It consists of two part, namely

- the ACTUS Data Dictionary (ADD) and

- the ACTUS Cashflow Generation Algorithm (ACGA).

The first standardizes the data involved in a financial contract and the second
standardizes the definition of financial contracts, i.e., the rules that define 
who cashflows are exchanged by the involved parties.




## Most elementary building blocks: financial contracts

The standard for both parts is defined in a way that it incorporates essential
parts of the basic banking knowledge, and by this supporting automation across the
board in a particular way. The reason ACTUS is focussing on financial contracts 
is that such contracts are the basic building blocs of all financial business
and thus the technical lynchpin of the financial system. Indeed, the most basic 
action in all financial activity is the individual single cashflow.
As, e.g., if you put money on your current account or withdraw money from it.
Usually, however, cashflows do not occur individually but are tied together by
some rules. As, e.g., in the case of a bond or a mortgage. And these rules define
the financial contract. 

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width = 15cm]{./figures/bowtie.pdf}
  \end{center}
  \caption{Bowtie structure of the financial system with the ACTUS contract types
  at its center.}
  \label{fig:Bowtie}
\end{figure}
```

Financial contracts have a central role in the structure
of the financial system. This is illustrated in figure \ref{fig:Bowtie}. This 
structure is often found in complex systems showing up in nature and society. 
An example outside the world of finance is the metabolism. 
It contains very diverse inputs (left) and outputs (right) but a highly standardized
and efficient core in the center. 



[THIS HAS TO BE ELABORATED ON SOMEWHAT MORE.]

\clearpage

## The logic of ACTUS

The logic of the processing of ACTUS contracts is shown in figure \ref{fig:ACTUS-logic}.

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width = 15cm]{./figures/ACTUS_logic.pdf}
  \end{center}
  \caption{The processing logic of ACTUS contracts.}
  \label{fig:ACTUS-logic}
\end{figure}
```


\clearpage

## Aim of the FEMSdevPkg package

Notice that ACTUS is an infrastructural element that cannot be used in isolation
but needs an environment to make use of it.

that has the power to 
overcome this difficulty and provides the means to become aquainted with this
layer and use it in a simple way. 


The \verb+FEMSdevPkg+ is an R package that give users the possibility to try out
the modeling of single financial contracts, portfolios of contracts or whole 
institution on a demo level. The package uses the ACTUS contract standard and 
cash flow generating algorithms that are transparent, efficient and meant to be
usable for even large scale institutions. The aim of this package, however. is
to give users an introduction an easy access to this technology at a demonstration 
level. For production level systems the interest users should look for software
solution based on the ACTUS technology.\footnote{%
On exsample is the product suite of the Ariadne Business Analytics AG.}

\newpage

# Setup
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pandoc)
```

```{r setup, warning = FALSE}
library(FEMSdevPkg)
```
For the calculations, a server from [ACTUS Financial Research Foundation](www.actusfrf.com) hosted by ZHAW is used.
Here we define the serverURL required for all calculations.
```{r}
serverURL <- "https://dadfir3-app.zhaw.ch/"
```
Now we are all set to start creating contracts.

# Single Contracts
## Fixed Rate Contracts
Let's start with a simple PAM (Principal at Maturity) contract.
```{r}
pam1 <- bondvr("2013-12-31", maturity = "5 years", nominal = 1000,
               coupon = 0.02, paymentFreq = "1 year", role = "long",
               rateResetFreq = "Fixed rate")
```

We can have a look at the contract terms.
```{r}
unlist(pam1$contractTerms)
```

With the function `generateEventSeries()` we can generate all events of the contract.
The functions takes the following arguments:
- `contract`: The contract object
- `riskFactors``: A list of risk factors (can also be empty)
- `serverURL`: The URL of the server to be used for the calculations

```{r}
evs1 <- generateEventSeries(pam1, list(), serverURL)
unlist(list(contractID = evs1$contractID,
            contractType=evs1$contractType,
            statusDate= evs1$statusDate,
            riskFactors = evs1$riskFactors
))
evs1$events_df
```

The events can be plotted afterwards by calling the function `cashflowPlot()`.
```{r}
cashflowPlot(evs1)
```

## Variable Rate Contracts
Contracts can also have variable interest rates. Herefore, we need to initialize some interest rate risk factors first.

First of all, let's load the sample data from the package.
Suggestion: We could make it available inside the package and load it from there.
```{r}
mydatadir <- "./mydata"
installSampleData(mydatadir)
```

Now we can load the sample data.
```{r}
falling_fp <- paste0(mydatadir,"/UST5Y_fallingRates.csv")
rising_fp <-  paste0(mydatadir,"/UST5Y_risingRates.csv")
rfx_falling <- sampleReferenceIndex(falling_fp,"UST5Y_fallingRates", 
                                    "YC_EA_AAA",100)
rfx_rising <- sampleReferenceIndex(rising_fp,"UST5Y_risingRates",
                                   "YC_EA_AAA",100)
```

After that we can create a contract with variable rates and generate the events and outputs as before.
```{r}
pam2 <- bondvr("2020-12-31", maturity = "5 years", nominal = 50000,
               coupon = 0.02, paymentFreq = "3 months", role = "long",
               rateResetFreq = "1 years", rateResetSpread = 0.01 )
unlist(pam2$contractTerms)
evs2 <- generateEventSeries(pam2, list(rfx_falling), serverURL)
evs2$events_df
```
And also the cashflow plot.
**Need to figure out a way to adjust the size of the plot, not good yet!!**
```{r, width = 12, height = 8}
cashflowPlot(evs2)
```


## Including raw Latex

Here we have a simple exmaple of a small table written in Latex

```{=latex}
\begin{tabular}{ll}
A & B \\
A & B \\
\end{tabular}
```


## Examples from Henriettes FEMS documentation:
Start with an Example picture to see whether referencing to figures works (picture replaced with EXAMPLE).

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width = 4cm]{./figures/example.jpg}
  \end{center}
  \caption{Structure of balance sheet and profit and loss statement of a very simple model bank.}
  \label{fig:BankBBsimple}
\end{figure}
```

Text which is formatted in latex has to wrapped in a block with the {=latex} command:
References etc. do not need to be changed 

```{=latex}
The second important part of the accounting information is the P\&L statement shown in the lower part of 
Fig.~\ref{fig:BankBBsimple}. It reports all the revenues and expenses during the reporting period 
(typically a year or a quarter). For our model bank we have two types of revenues and two types of expenses:
\begin{itemize}
  \item 
  Expenses:
  \begin{itemize}
    \item 
    Generated out of the balance sheet:\\ 
    Interest payments and expenses due to change of the valuation.
    \item
    Expenses for running the business:\\
    Wages, rents, IT
  \end{itemize}
  \item
  Revenues:
  \begin{itemize}
    \item 
    Generated out of the balance sheet:\\
    Interest income and capital gains.
    \item
    Additional revenues from running the business, mainly fees for services.
  \end{itemize}
\end{itemize}
In the following we will concentrate on modeling the balance sheet and the expenses and revenues generated
out of the balance sheet in quite some details, while the additional expenses and revenues are taken into 
account summarily.
```


## Example for a formula:
Here we see an example for a formula which even uses R values within.
If we include R values in markdown we have to change the /Sexpr{} from latex (sweave) with the expression: 'r inline_code_here'

```{r ir, echo =FALSE}
#Value which is used in formula notation has to be defined in r environment
ir = 0.04
```

```{=latex}
Determining a contract's value is refered to as valuation. 
The value of a contract is in general not simply given, it requires computations.  
The type of computation is refered to as valuation method. 
In accounting there are different valuation methods. 
In this course, we will only use two of them, namely \emph{Nominal Value} and \emph{Marking to Model}.
\begin{description}
  \item[Nominal value]
  The value of the contract is set to its nominal value. 
  This is the simplest valuation method which only requires extracting this value from the 
  contract data.
  \item[Marking to Model]
  The value of the contract is determined by means of a mathematical model. 
  In the case of a bond or a loan, the value is determined by its net present value computed 
  through discounting. In cases without default risk, as is the case e.g. for Swiss government bonds, 
  the risk-free interest rate is used for discounting. 
  If the market interest rate (for all terms, to keep it simple) is 3\%, the bond's net present
  value is
  \begin{equation}
  \label{eq:NPV_bond}
    \text{NPV } = \sum_{i=1}^{5} \frac{40}{`r 1+ir`^i} + \frac{1000}{`r 1+ir`^5} = 
    `r round(sum(rep(40,5)/(1+ir)^(1:5)) + 1000/(1+ir)^5,2)`
  \end{equation}
\end{description}
In the simulation environment, valuation is done with the command \texttt{value(object, type, method)}.
```

## Including a figure with Latex (Picture replaced with EXAMPLE picture):

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width=10cm]{./figures/example.jpg}
  \end{center}
  \caption{Time dependent net present value of the example bond for different market interest rates. 
  Full black line: 4\%; dashed red line: Stress scenario 1, 3\%; 
  dashed green line: Stress scenario 2, 5\%.}
  \label{fig:NPV_bond}
\end{figure}
```


Example for a pagebreak:
\pagebreak

```{=latex}
This is how to cross-reference: Cf. fig. \ref{fig:NPV_bond}.
```

## Including a table with R values
First we have an  example for a simple table in latex. No changes to the latex 
code are needed. Afterwards we see another formula. Again the format of the 
inline r code had to be changed to 'r inline_code_here'
```{r values for tabular block, echo = FALSE}
#Values which are used in formula notation have to be defined in r environment
val <- 1000
val.lo <- 800
val.hi <- 1200
```


```{=latex}
There is a theory to compute this value, which is closely related to a quantity called 
the \emph{duration} of a bond. 
Here, we will neither introduce these notions nor do the theoretical computation. 
This is part of the module \emph{Mathematik der Finanzmärkte 1}. 
Instead, we will carry out an approximate numerical computation of this quantity using the bond value 
obtained for the different values of the market interest rate.

The values obtained from the simulation were:
\begin{center}
  \begin{tabular}{c|ccc}
  Interest rate & 3\% & 4\% & 5\% \\\hline
  Value & `r val.hi` &`r val` & `r val.lo`
  \\
  \end{tabular}
\end{center}
The bond's sensitivity is thus approximately given as 
\[
  \frac{1}{`r val`}\cdot \frac{`r val.hi`- `r val`}{0.01} = `r (val.hi-val)/(0.01*val)`
\]
This value is pretty close to the bonds term to maturity, which is 5 years. This is not accidental. 
Indeed, a bonds sensitivity is the negative value of it's duration, and the duration is closely tied 
to the bond's term to maturity.\footnote{Both quatities are equal for a zero-coupon bond. 
For a coupon-bearing the duration is strictly smaller than the term to maturity, 
and the difference is larger the larger the market interest rate.}

```



