---
title: "portfolio"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{portfolio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: "Henriette Elise Breymann, Gian-Andrin Tommasini"
date: "`r Sys.Date()`"
---
# Setup
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pandoc)
```

```{r setup, warning = FALSE}
library(FEMSdevPkg)
```
For the calculations, a server from [ACTUS Financial Research Foundation](www.actusfrf.com) hosted by ZHAW is used.
Here we define the serverURL required for all calculations.
```{r}
serverURL <- "https://dadfir3-app.zhaw.ch/"
```
Now we are all set to start creating contracts.

# Single Contracts
## Fixed Rate Contracts
Let's start with a simple PAM (Principal at Maturity) contract.
```{r}
pam1 <- bondvr("2013-12-31", maturity = "5 years", nominal = 1000,
               coupon = 0.02, paymentFreq = "1 year", role = "long",
               rateResetFreq = "Fixed rate")
```

We can have a look at the contract terms.
```{r}
unlist(pam1$contractTerms)
```

With the function `generateEventSeries()` we can generate all events of the contract.
The functions takes the following arguments:
- `contract`: The contract object
- `riskFactors``: A list of risk factors (can also be empty)
- `serverURL`: The URL of the server to be used for the calculations

```{r}
evs1 <- generateEventSeries(pam1, list(), serverURL)
unlist(list(contractID = evs1$contractID,
            contractType=evs1$contractType,
            statusDate= evs1$statusDate,
            riskFactors = evs1$riskFactors
))
evs1$events_df
```

The events can be plotted afterwards by calling the function `cashflowPlot()`.
```{r}
cashflowPlot(evs1)
```

## Variable Rate Contracts
Contracts can also have variable interest rates. Herefore, we need to initialize some interest rate risk factors first.

First of all, let's load the sample data from the package.
Suggestion: We could make it available inside the package and load it from there.
```{r}
mydatadir <- "./mydata"
installSampleData(mydatadir)
```

Now we can load the sample data.
```{r}
falling_fp <- paste0(mydatadir,"/UST5Y_fallingRates.csv")
rising_fp <-  paste0(mydatadir,"/UST5Y_risingRates.csv")
rfx_falling <- sampleReferenceIndex(falling_fp,"UST5Y_fallingRates", 
                                    "YC_EA_AAA",100)
rfx_rising <- sampleReferenceIndex(rising_fp,"UST5Y_risingRates",
                                   "YC_EA_AAA",100)
```

After that we can create a contract with variable rates and generate the events and outputs as before.
```{r}
pam2 <- bondvr("2020-12-31", maturity = "5 years", nominal = 50000,
               coupon = 0.02, paymentFreq = "3 months", role = "long",
               rateResetFreq = "1 years", rateResetSpread = 0.01 )
unlist(pam2$contractTerms)
evs2 <- generateEventSeries(pam2, list(rfx_falling), serverURL)
evs2$events_df
```
And also the cashflow plot.
**Need to figure out a way to adjust the size of the plot, not good yet!!**
```{r, width = 12, height = 8}
cashflowPlot(evs2)
```


## Including raw Latex

Here we have a simple exmaple of a small table written in Latex

```{=latex}
\begin{tabular}{ll}
A & B \\
A & B \\
\end{tabular}
```


## Examples from Henriettes FEMS documentation:
Start with an Example picture to see whether referencing to figures works (picture replaced with EXAMPLE).

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width = 4cm]{./figures/example.jpg}
  \end{center}
  \caption{Structure of balance sheet and profit and loss statement of a very simple model bank.}
  \label{fig:BankBBsimple}
\end{figure}
```

Text which is formatted in latex has to wrapped in a block with the {=latex} command:
References etc. do not need to be changed 

```{=latex}
The second important part of the accounting information is the P\&L statement shown in the lower part of 
Fig.~\ref{fig:BankBBsimple}. It reports all the revenues and expenses during the reporting period 
(typically a year or a quarter). For our model bank we have two types of revenues and two types of expenses:
\begin{itemize}
  \item 
  Expenses:
  \begin{itemize}
    \item 
    Generated out of the balance sheet:\\ 
    Interest payments and expenses due to change of the valuation.
    \item
    Expenses for running the business:\\
    Wages, rents, IT
  \end{itemize}
  \item
  Revenues:
  \begin{itemize}
    \item 
    Generated out of the balance sheet:\\
    Interest income and capital gains.
    \item
    Additional revenues from running the business, mainly fees for services.
  \end{itemize}
\end{itemize}
In the following we will concentrate on modeling the balance sheet and the expenses and revenues generated
out of the balance sheet in quite some details, while the additional expenses and revenues are taken into 
account summarily.
```


## Example for a formula:
Here we see an example for a formula which even uses R values within.
If we include R values in markdown we have to change the /Sexpr{} from latex (sweave) with the expression: 'r inline_code_here'

```{r ir, echo =FALSE}
#Value which is used in formula notation has to be defined in r environment
ir = 0.04
```

```{=latex}
Determining a contract's value is refered to as valuation. 
The value of a contract is in general not simply given, it requires computations.  
The type of computation is refered to as valuation method. 
In accounting there are different valuation methods. 
In this course, we will only use two of them, namely \emph{Nominal Value} and \emph{Marking to Model}.
\begin{description}
  \item[Nominal value]
  The value of the contract is set to its nominal value. 
  This is the simplest valuation method which only requires extracting this value from the 
  contract data.
  \item[Marking to Model]
  The value of the contract is determined by means of a mathematical model. 
  In the case of a bond or a loan, the value is determined by its net present value computed 
  through discounting. In cases without default risk, as is the case e.g. for Swiss government bonds, 
  the risk-free interest rate is used for discounting. 
  If the market interest rate (for all terms, to keep it simple) is 3\%, the bond's net present
  value is
  \begin{equation}
  \label{eq:NPV_bond}
    \text{NPV } = \sum_{i=1}^{5} \frac{40}{`r 1+ir`^i} + \frac{1000}{`r 1+ir`^5} = 
    `r round(sum(rep(40,5)/(1+ir)^(1:5)) + 1000/(1+ir)^5,2)`
  \end{equation}
\end{description}
In the simulation environment, valuation is done with the command \texttt{value(object, type, method)}.
```

## Including a figure with Latex (Picture replaced with EXAMPLE picture):

```{=latex}
\begin{figure}[h] 
  \begin{center}
  \includegraphics[width=10cm]{./figures/example.jpg}
  \end{center}
  \caption{Time dependent net present value of the example bond for different market interest rates. 
  Full black line: 4\%; dashed red line: Stress scenario 1, 3\%; 
  dashed green line: Stress scenario 2, 5\%.}
  \label{fig:NPV_bond}
\end{figure}
```

Example for a pagebreak:
\newpage

## Including a table with R values
First we have an  example for a simple table in latex. No changes to the latex code are needed.
Afterwards we see another formula. Again the format of the inline r code had to be chagned to 'r inline_code_here'
```{r values for tabular block, echo = FALSE}
#Values which are used in formula notation have to be defined in r environment
val <- 1000
val.lo <- 800
val.hi <- 1200
```


```{=latex}
There is a theory to compute this value, which is closely related to a quantity called 
the \emph{duration} of a bond. 
Here, we will neither introduce these notions nor do the theoretical computation. 
This is part of the module \emph{Mathematik der Finanzmärkte 1}. 
Instead, we will carry out an approximate numerical computation of this quantity using the bond value 
obtained for the different values of the market interest rate.

The values obtained from the simulation were:
\begin{center}
  \begin{tabular}{c|ccc}
  Interest rate & 3\% & 4\% & 5\% \\\hline
  Value & `r val.hi` &`r val` & `r val.lo`
  \\
  \end{tabular}
\end{center}
The bond's sensitivity is thus approximately given as 
\[
  \frac{1}{`r val`}\cdot \frac{`r val.hi`- `r val`}{0.01} = `r (val.hi-val)/(0.01*val)`
\]
This value is pretty close to the bonds term to maturity, which is 5 years. This is not accidental. 
Indeed, a bonds sensitivity is the negative value of it's duration, and the duration is closely tied 
to the bond's term to maturity.\footnote{Both quatities are equal for a zero-coupon bond. 
For a coupon-bearing the duration is strictly smaller than the term to maturity, 
and the difference is larger the larger the market interest rate.}

```



